# Codebook for the assignment

The following Codebook was created on 2017-11-16. Please review analysis.Rmd or run_analysi.html for details. 

## Variable list and descriptions

I am using R function to list the name of variable and description

```{r}
library(knitr)
read.table(text = "Name of variable | Description
featDomain       | Time domain signal or frequency domain signal (Time or Freq)
featInstrument   | Measuring instrument (Accelerometer or Gyroscope)
featAcceleration | Acceleration signal (Body or Gravity)
featVariable     | Variable (Mean or SD)
featJerk         | Jerk signal
featMagnitude    | Magnitude of the signals calculated using the Euclidean norm
featAxis         | 3-axial signals in the X, Y and Z directions (X, Y, or Z)
featCount        | Count of data points used to compute `average`
featAverage      | Average of each variable for each activity and each subject
subject          | ID the subject who performed the activity for each window sample. Its range is from 1 to 30.
activity         | Activity name", header = T, sep = "|") %>% 
  knitr::kable(col.names = c("Variable name", "Description"))
```

## View the structure of dataset

```{r}
str(tidy_data)
```

## The key variables in the data table

```{r}
key(tidy_data)
```

## Review some of the data

```{r}
tidy_data
```

## View all the feature combinations using grep function

```{r}
tidy_data[, .N, by=c(names(tidy_data)[grep("^feat", names(tidy_data))])]
```

## Now let's look at summary statistics of variables

```{r}
# View summary of datatable
summary(tidy_data)
```


## Save the data to a text file. Step 5 of the instruction.
------------

I am going to save the data table object to a tab-delimited text file called `humanactivity_recognition_Smartphones.txt`. 
The code is as follows:

```{r save}
write.table(x = tidy_data, file = "humanactivity_recognition_Smartphones.txt", quote=FALSE, sep="\t", row.names=FALSE)
```
